# Simulation and Profiling

## Str
str function shows structure
very useful

```{r}
library(datasets)
s <- split(airquality, airquality$Month)
str(s)
```

## Simulation

### Generating Random Numbers

* `rnorm` - generate random Normal variates with a given mean and standard deviation
* `dnorm` - evalute the Normal probability density (with a given mean/SD) at a point (or vector of points)
* `pnorm` - evaluate the cumulative distribution function for a Normal distribution
* `rpois` - random Poisson variates with a given rate

Probability distribution functions usually have four functions. By prefix:
- d: density
- r: random number generation
- p: cumulative distribution
- q: quantile function

```{r} 
# make sure to set seed if you need reproducibility! - which you will during 
# simulation
set.seed(1)
x <- rnorm(10)
x
x <- rnorm(10, 20, 2)
x
summary(x)
```

#### Poisson
```{r}

rpois(10, 1)

ppois(2, 2)
```

### Simulating a Linear Model

#### From a normal distribution
```{r}
set.seed(20)
x <- rnorm(100)
e <- rnorm(100, 0, 2)

y <- 0.5 + 2 * x + e
summary(y)
plot(x, y)
```

#### From a binomial distribution
```{r}
set.seed(20)
x <- rbinom(100, 1, 0.5)
e <- rnorm(100, 0, 2)

y <- 0.5 + 2 * x + e
summary(y)
plot(x, y)
```

#### From a Poisson distribution
```{r}
set.seed(1)
x <- rnorm(100)
log.mu <- 0.5 + 0.3 * x
y <- rpois(100, exp(log.mu))
summary(y)
plot(x, y)
```

### Sampling
```{r}
set.seed(1)
sample(1:10, 4)

sample(1:10)
sample(1:10, replace = TRUE) # items can be repeated
```

## Profiling

can use `system.time`, just returns user (cpu) time and elapsed time

can use `Rprof` and `summaryRprof` for more detailed output

### summaryRprof

`$by.total` - total time
`$by.self` - time of function subtracting child functions in stack

Note: C or Fortran code not profiled
