---
title: "Loop functions"
output: html_document
---

These functions are good as shortcuts, especially on the command line for
exploratory work, so you don't have to write a loop.

`lapply` is like a map
```{r}
x <- list(a = 1:5, b = rnorm(1))
lapply(x, mean)

x <- 1:4
lapply(x, runif, min = 0, max = 10)

x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x
lapply(x, function(elt) elt[, 1]) # returns first column of each matrix
```

`sapply` is like `lapply` but it tries to simplify the list
```{r}
x <- list(a = 1:5, b = rnorm(1))
simplified <- sapply(x, mean)
simplified
class(simplified)
```

`apply` is *a bit* like reduce - it can summarize/collapse a dimension

```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 2, mean) # collapses the rows to get the mean of each column
apply(x, 1, mean) # collapses the columns to get the mean of each row
```

**NOTE:** better to use `rowMeans` and `colMeans` for the above, as they are
optimized

```{r} 
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))

```

```{r} 
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
apply(a, c(1, 2), mean)
rowMeans(a, dims = 2)
```

`mapply` is a "multivariate" version of apply. It allows you to apply a function
over multiple arguments in parallel, and they can refer to one another.

```{r}
mapply(rep, 1:4, 4:1)
```

```{r}
noise <- function(n, mean, sd) {
  rnorm(n, mean, sd)
}
# using mapply you can vectorize non-vector functions
mapply(noise, 1:5, 1:5, 2)
```

`tapply` is applies function over subsets of a vector

```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
f

tapply(x, f, mean)
tapply(x, f, mean, simplify = FALSE)
tapply(x, f, range)
```

`split` not a loop function - but useful to know in this context
splits vector into groups - like `partition` in a Laravel collection
```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)

split(x, f)

# common to use split followed by lapply - but you could have used tapply
lapply(split(x, f), mean)

library(datasets)
head(airquality)

s <- split(airquality, airquality$Month)
lapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
```

## Splitting on more than one level

```{r}
x <- rnorm(10)
f1 <- gl(2, 5)
f2 <- gl(5, 2)

f1
f2

interaction(f1, f2)
str(split(x, list(f1, f2), drop = TRUE))
```
